---
title: "p8105_hw5_jc4166"
author: "Jerri Chen"
date: "11/08/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(rvest)

```

## Problem 1
```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

Write a function to replace the missing values per the following guidelines:  
-For numeric variables, you should fill in missing values with the mean of non-missing values  
-For character variables, you should fill in missing values with "virginica"  

```{r}
missing_data = function(x) {
  output = vector(length = length(x))
  
  for (i in 1:length(x)) {
    
    if (is.numeric(x[i])) {
      mean_i = mean(x[!is.na(x)])
          if (!is.na(x[i])) {output[i] = x[i]} 
        else {output[i] = mean_i}
    }
    else if (is.character(x[i])) {
        if (!is.na(x[i])) {output[i] = x[i]} 
      else {output[i] = "virginica"}
    }
  }
  output
}

iris_with_missing =  map(iris_with_missing, missing_data) %>% 
  as_tibble()
```


## Problem 2
```{r, message=FALSE, warning=FALSE}
all_files = list.files("./data")

all_data = tibble(file_name = all_files) %>% 
  mutate(contents = map(file_name, ~read_csv(file.path("./data", .)))) %>% 
  unnest() %>% 
  mutate(subject_id = file_name) %>% 
  select(subject_id, everything()) %>% 
  separate("file_name", into = c("arm")) %>% 
  mutate(
    arm = recode(arm, 
                      "con" = "control",
                      "exp" = "experimental"),
    subject_id = substr(subject_id, 1, 6)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "observation"
  ) %>% 
  mutate(
    week = substr(week, 6, 7),
    week = as.numeric(week))
```
The original data came as separate csv files for each participant. The data were combined into a single data frame, and tidied such that each variable is a single column (study arm, week of study, and observed value). Of note, if I wanted to present the data in a table to a human reader, I would not have used pivot_longer as this format is harder to read. 

Make a spaghetti plot showing observations on each subject over time
```{r}
all_data %>% 
  ggplot(aes(x = week, y = observation, color = subject_id)) + 
  geom_path() + 
  facet_grid(~arm) +
  labs(
    title = "Observations for Participants Over Time",
    x = "Week",
    y = "Observation"
   ) + 
  theme(legend.position = "bottom")
```

Comparing the control to the experimental participants, the observations for the control group do not appear to change over time, whereas the observations for the experimnetal group increase over time.

# Problem 3
Conduct a simulation to explore power in a simple linear regression.
```{r}
sim_regression = function(n = 30, beta0 = 2, beta1) {
  sim_data = tibble(
    x = rnorm(n, mean = 1, sd = 1),
    y = beta0 + beta1 * x + rnorm(n, 0, 50)
  )
  
ls_fit = lm(y ~ x, data = sim_data) %>% 
  broom::tidy()

}
```

Set β1=0. Generate 10000 datasets from the model.
```{r}
sim_results = 
  rerun(10000, sim_regression(30, 2, 0)) %>% 
  bind_rows()

head(sim_results)
```

Repeat the above for β1={1,2,3,4,5,6}.
```{r, eval=FALSE}
sim_results_repeat = 
  tibble(b1 = c(1,2,3,4,5,6)) %>% 
  mutate(
    output_list = map(.x = b1, ~rerun(10000, sim_regression(beta1 = .x))),
    estimate_dfs = map(output_list, bind_rows)
  ) %>% 
  select(-output_list) %>% 
  unnest(estimate_dfs) %>% 
  mutate(
    reject_null = ifelse(p_value < alpha, 1, 0)
  )
```


Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of β1 on the x axis. Describe the association between effect size and power.
